\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{amsfonts,amssymb,amsmath}

\begin{document}

\title{Constraining gravitational wave amplitude birefringence with GWTC-3}

\author{Thomas C.K. Ng}

\begin{abstract}
    
\end{abstract}

\section{Everything}
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GW) amplitude birefringence is a property of space, which GR does not included.

GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves, which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
In GR, the amplitude ratio of the left and right polarisations only depends on $\iota$, the angle between our line of sight and the orbital angular momentum of the binary black holes (BBH).
\begin{equation}
    \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}where $h_L$ and $h_R$ are the amplitude of left and right polarizations of the GW respectively.
For the birefringence property mentioned in \citet{Maria_2021} and other papers, the amplitude ratio depends on not only $\iota$, but also the distance to the merger and the strength of the birefringence.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
\end{equation}where $d_C$ is the comoving distance to the merger, $\widetilde{\kappa}$ is the opacity parameter that represent the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are the observed amplitude of left and right polarizations of the GW respectively.
This is a preliminary version of the birefringence model in CS, which assume there is no frequency dependence.
This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

In this paper, we considered the frequency dependence of the birefringence as well. Thus, the amplitude ratio of the left and right polarisations will be as follow.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{new}=\frac{\exp\left({\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}}\right)\left(1-\cos\iota\right)^2}{\exp\left({-\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}}\right)\left(1+\cos\iota\right)^2}\,,
\end{equation}where $f$ is the frequency of the GW and $\kappa$ is the dimensionless opacity parameter that represent the strength of the birefringence.
This frequency dependence provides stronger enhancement or suppression from the birefrigence to high frequency components of the GW, and visa versa.
Also, the frequency dependence can break the degeneracy between kappa and iota, as this would make the effect of the birefringence affect the amplitude ratio differently compared to iota.

We performed estimations on $\kappa$ using Bilby with data from the GWTC catalog to constrain the plausible range of $\kappa$.
Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from interferometer and priors of the parameters provided. \citep{Ashton_2019}

We assume the GW are produced at the BBH following GR. As the GW propagate through space, the effect of the birefringence will be built up as the distance increases.
To implement the amplitude birefringence on Bilby, waveforms are modified according to the following equation:
\begin{equation}
    h_\mathrm{L,R}^{\mathrm{Biref}}=
    h_\mathrm{L,R}^{\mathrm{GR}}\times
    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{ Gpc}}\frac{f}{100\mathrm{ Hz}}\right)\,.
\end{equation}
This modification allows Bilby to perform estimations on $\kappa$, as the waveforms will depends on $\kappa$ as well.

GWTC-3 Results

Limitation\\
orientation
sensitivity

Future Study\\
Data from more detectors

\section{Introduction}



\section{Method}



\section{Results}



\section{Discussion}



\section{Acknowledgements}



\bibliography{bib}

\end{document}